### 2014-10-17 ApTranscriptome_TR.Rmd
working on figure for genetic assimilation

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

# Compare expression at optimum temp (A22.opt) between colonies using t-test
t.test(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt)
boxplot(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt)

# T test on log-transformed values to control for outliers
t.test(log(A22.high.transcripts$A22.opt+1), log(A22.high.transcripts$Ar.opt+1))
boxplot(log(A22.high.transcripts$A22.opt+1), log(A22.high.transcripts$Ar.opt+1))
```

@@

```{r A22_high_wilcoxon}
w1 <- wilcox.test(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w1
```

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), ]

# t-test with log values
t.test(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))
boxplot(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))

# Wilcoxon signed rank-test
w2 <- wilcox.test(Ar.low.transcripts$A22.opt, Ar.low.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w2
```

@@

```{r induction_vs_constitutive}
# calculate degree of induction
Ap.induction <- ddply(resp.TPM.dt.sub.pred, .(colony, Transcript), summarise, 
                      upreg = (max(pTPM) - min(pTPM))/min(pTPM) * 100 )
# merge optimum expression, in Ap.response.type data.frame, with predicted TPM data
Ap.induction <- data.table(Ap.induction)
setkey(Ap.induction, Transcript)
Ap.response.type2 <- Ap.induction[Ap.response.type]

# plot constitutive expression vs induction for High, Low and Bimodal genes only

setkey(Ap.response.type2, A22.type)
Ap.response.type2.up <- Ap.response.type2[A22.type!="Intermediate", ]
setkey(Ap.response.type2.up, Ar.type)
Ap.response.type2.up <- Ap.response.type2.up[Ar.type!="Intermediate", ]

gg5 <- ggplot(Ap.response.type2.up, aes(x = log(A22.opt), y = log(upreg), colour = A22.type)) + 
  geom_point() + 
  facet_grid(. ~ colony)

print(gg5)
```

>>>>

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts.df <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

# Compare expression at optimum temp (A22.opt) between colonies using t-test
t.test(A22.high.transcripts.df$A22.opt, A22.high.transcripts.df$Ar.opt)
boxplot(A22.high.transcripts.df$A22.opt, A22.high.transcripts.df$Ar.opt)

# T test on log-transformed values to control for outliers
t.test(log(A22.high.transcripts.df$A22.opt+1), log(A22.high.transcripts.df$Ar.opt+1))
boxplot(log(A22.high.transcripts.df$A22.opt+1), log(A22.high.transcripts.df$Ar.opt+1))
```

@@

```{r A22_high_wilcoxon}
w1 <- wilcox.test(A22.high.transcripts.df$A22.opt, A22.high.transcripts.df$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w1
```

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low.transcripts.df <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), ]

# t-test with log values
t.test(log(Ar.low.transcripts.df$A22.opt+1), log(Ar.low.transcripts.df$Ar.opt+1))
boxplot(log(Ar.low.transcripts.df$A22.opt+1), log(Ar.low.transcripts.df$Ar.opt+1))

# Wilcoxon signed rank-test
w2 <- wilcox.test(Ar.low.transcripts.df$A22.opt, Ar.low.transcripts.df$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w2
```

@@

```{r induction_vs_constitutive}
# calculate degree of induction
Ap.induction <- ddply(resp.TPM.dt.sub.pred, .(colony, Transcript), summarise, 
                      upreg = (max(pTPM) - min(pTPM))/min(pTPM) * 100 )
# cast to wide
Ap.induction.wide <- dcast(Ap.induction, Transcript ~ colony, value.var = "upreg")
names(Ap.induction.wide)[names(Ap.induction.wide)=="A22"] <- "A22.upreg"
names(Ap.induction.wide)[names(Ap.induction.wide)=="Ar"] <- "Ar.upreg"

# merge optimum expression, in Ap.response.type data.frame, with predicted TPM data
Ap.induction.wide <- data.table(Ap.induction.wide)
setkey(Ap.induction.wide, Transcript)
Ap.response.type2 <- Ap.induction.wide[Ap.response.type]

# plot constitutive expression vs induction for High, Low and Bimodal genes only
setkey(Ap.response.type2, A22.type)
up.terms <- c("Bimodal", "Low", "High")
Ap.response.type2.up <- Ap.response.type2[A22.type %in% up.terms, ]

setkey(Ap.response.type2.up, Ar.type)
Ap.response.type2.up <- Ap.response.type2.up[Ar.type %in% up.terms, ]
Ap.response.type2.up <- droplevels(Ap.response.type2.up)

gg5 <- ggplot(Ap.response.type2.up, aes(x = log(Ar.opt), y = log(A22.upreg), colour = A22.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. carolinensis)", y = "log(Degree of inducibility in A. picea)") +
  theme(legend.position = "none")

gg6 <- ggplot(Ap.response.type2.up, aes(x = log(A22.opt), y = log(Ar.upreg), colour = Ar.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. picea)", y = "log(Degree of inducibility in A. carolinensis)") +
  theme(legend.justification=c(1,0), legend.position=c(1,0), legend.text = element_text(size = 12))

# grid.arrange
tiff("results/constitutive_vs_inducibility.tiff", compression = "lzw", bg = "white")
grid.arrange(gg5, gg6, nrow = 1)
dev.off()

### plot constitutive expression in one species vs the other
gg7 <- ggplot(Ap.response.type2.up, aes(x = log(Ar.opt), y = log(A22.opt), colour = A22.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. carolinensis)", y = "log(constitutive expression in A. picea)") +
  theme(legend.position = "none")  
gg7

### plot inducibility in one species vs the other
gg8 <- ggplot(Ap.response.type2.up, aes(x = log(Ar.upreg), y = log(A22.upreg), colour = A22.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(upregulation in A. carolinensis)", y = "log(upregulation in A. picea)") +
  theme(legend.position = "none")  
gg8

### constitutive vs upregulated for *High* genes only
Ap.response.type2.A22.high <- Ap.response.type2[Ap.response.type2$Transcript %in% A22.high.transcripts.df$Transcript, ]
str(Ap.response.type2.A22.high)

gg9 <- ggplot(Ap.response.type2.A22.high, aes(x = log(Ar.opt), y = log(A22.opt), colour = Ar.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. carolinensis)", y = "log(Degree of inducibility in A. picea)") +
  theme(legend.position = "none")  
gg9
